--- 
params:
    group: "xan_test"
    refs: "22_331_assembly"
    samp_data: "_test_data/metadata_medium.csv"
    ref_data: "_test_data/merged_assembly_stats.tsv"
    sendsketch: "_test_data/sendsketch"
    variant_data: "_test_data/variant_data"
    ani_matrix: "_test_data/xan_test_comp.csv"
    core_phylo: "_test_data/xan_test.treefile"
    multiqc: "_test_data/multiqc"
    quast: "_test_data/quast"
    versions: "_test_data/software_versions.yml"
execute:
  echo: false
---

# Summary

```{r knitr_settings, file='knitr_settings.R'}
```

```{r load_libraries, warning=FALSE, message=FALSE, file='load_libraries.R'}
# When running manually for development: source('load_libraries.R')
```

```{r parse_inputs, file='parse_inputs.R'}
# When running manually for development: source('parse_inputs.R')
```

This report is produced by the **`nf-core` `pathogensurveillance` pipeline** version **{{< var version >}}**.

| <!-- --> | <!-- --> | 
|-----|-----|
| **Date**            | TBD |
| **Last updated**    | `r format(Sys.time(), '%B %d , %Y')` |
| **Submitter(s):**   | `r combine_words(unique(samp_data$report_group))` |
| **Samples:**        | `r nrow(samp_data)` |

## Run info

This is the first page a user sees.
What should go here?

* status of pipeline execution
* table of run metadata (run_id, submitter, time)
* sample metadata ( sample ids, type of sequencing, input metadata)

Could copy the info that nextflow tower shows.

```{r}
DT::datatable(samp_data)
```

## Input settings

## Methods used

The `pathogen surveillance` pipeline used the following tools that should be referenced as appropriate:

  - Software versions used: `r xfun::embed_file('_test_data/software_versions.yml')`
  - A sample is first identified to genus using sendsketch and further identified to species using sourmash [@brown2016sourmash].
  - The `nextflow` data-driven computational pipeline enables deployment of complex parallel and reactive workflows [@di2017nextflow].

## About

The **`pathogen surveillance` pipeline** was developed by:

  - Zach Foster, Martha Sudermann, Camilo Parada-Rojas, Fernanda Iruegas-Bocardo, [Jeff Chang](http://changlab.cgrb.oregonstate.edu/) and [Nik Grunwald](http://grunwaldlab.cgrb.oregonstate.edu/).
  - Contributors include [Alex Weisberg](https://alexandra-weisberg.com/), ...
  
> To contribute or report bug please see our [github repository]( https://github.com/nf-core/pathogensurveillance). 
  
Please cite this pipeline in publications as follows: ...

