--- 
params:
    samp_data: "_test_data/metadata_medium.csv"
    ref_data: "_test_data/merged_assembly_stats.tsv"
    snp_phylos: "_test_data/xan_test_22_331_assembly.treefile"
    snp_align: "_test_data/xan_test_22_331_assembly.fasta"
    ani_matrix: "_test_data/comp.csv"
    core_phylo: "_test_data/xan_test.treefile"
execute:
  echo: false
---

```{r}
#| echo: false
#| warning: false
#| message: false
#| context: server
library(dplyr)
library(ggplot2)
library(readr)
```

# Summary {.unnumbered}

Here is your report produced by the `nf-core` `Pathogen Surveillance` pipeline **version {{< var version >}}**.

| <!-- --> | <!-- --> | 
|-----|-----|
| **Date**            | `r format(Sys.time(), '%B %d , %Y')` |
| **Submitter(s):**   | xxx |
| **Samples:**      | xxx |

## Run info

This is the first page a user sees.
What should go here?

* status of pipeline execution
* table of run metadata (run_id, submitter, time)
* sample metadata ( sample ids, type of sequencing, input metadata)

Could copy the info that nextflow tower shows.

```{r}
data <- read_csv('_test_data/metadata_medium.csv')
DT::datatable(data)
```

## Input settings

## Things that require user attention