--- 
params:
    samp_data: "_test_data/metadata_medium.csv"
    ref_data: "_test_data/merged_assembly_stats.tsv"
    snp_phylos: "_test_data/xan_test_22_331_assembly.treefile"
    snp_align: "_test_data/xan_test_22_331_assembly.fasta"
    ani_matrix: "_test_data/comp.csv"
    core_phylo: "_test_data/xan_test.treefile"
execute:
  echo: false
---

<!--# preloading libraries in one place at the top avoids mutliple reloads and metadata.-->
```{r}
#| echo: false
#| warning: false
#| message: false
#| context: server
library(dplyr)
library(ggplot2)
library(readr)
metadata <-read_csv('_test_data/metadata_medium.csv')
```

# Summary {.unnumbered}

This report is produced by the **`nf-core` `Pathogen Surveillance` pipeline** version **{{< var version >}}**.

| <!-- --> | <!-- --> | 
|-----|-----|
| **Date**            | TBD |
| **Last updated**    | `r format(Sys.time(), '%B %d , %Y')` |
| **Submitter(s):**   | `r unique(metadata$report_group)` |
| **Samples:**        | `r nrow(metadata)` |

## Run info

This is the first page a user sees.
What should go here?

* status of pipeline execution
* table of run metadata (run_id, submitter, time)
* sample metadata ( sample ids, type of sequencing, input metadata)

Could copy the info that nextflow tower shows.

```{r}
DT::datatable(metadata)
```

## Input settings

## Things that require user attention

## About

The **`nf-core` `Pathogen Surveillance` pipeline** was developped by:

  - Zach Foster and Nik Grunwald at USDA ARS, Corvallis, OR.
  - Martha Sudermann, Camilo Parada-Rojas, and Jeff Chang at Oregon State University, Corvallis, OR.

> To contribute or report bug please see our [github repository]( https://github.com/nf-core/pathogensurveillance). 
  

> Please cite this pipeline in publications as follows: ...

