---
title: Pathogen Surveillance Report
params:
  inputs: "_test_data/1"
execute:
  echo: false
---

```{r knitr_settings}
knitr::opts_chunk$set(echo = FALSE, fig.width = 10, warning = FALSE)
```

```{r load_libraries, warning=FALSE, message=FALSE}
# library(dplyr)
# library(ggplot2)
# library(readr)
# library(knitr)
# library(readr)
# library(purrr)
# library(yaml)
# library(phylocanvas)
# library(ape)
# library(magrittr)
# library(pheatmap)
# library(tidyverse)
# library(palmerpenguins)
# library(ade4)
# library(adegenet)
# library(poppr)
# library(ggtree)
# library(igraph)
# library(visNetwork)
# library(phangorn)
# library(ggplot2)
# library(ggnewscale)
# library(kableExtra)
# library(plotly)
# library(webshot2)
# library(metacoder)
```



## {{< ai doi >}} Bibliography

::: {#refs}
:::

# About

{{< bi file-earmark-person >}} The **`nf-core/pathogen surveillance` pipeline** was developed by: Zach Foster, Martha Sudermann, Camilo Parada-Rojas, Fernanda Iruegas-Bocardo, Ricardo Alcalá-Briseño, [Jeff Chang](http://changlab.cgrb.oregonstate.edu/) and [Nik Grunwald](http://grunwaldlab.cgrb.oregonstate.edu/).

Other contributors include: [Alex Weisberg](https://alexandra-weisberg.com/), ...

::: callout-tip
## Feedback

To contribute, provide feedback, or report bugs please visit our [github repository](https://github.com/nf-core/pathogensurveillance).
:::

Please cite this pipeline and nf-core in publications as follows:

{{< bi file-earmark-pdf-fill >}} Foster et al. 2024. PathogenSurveillance: A nf-core pipeline for rapid analysis of pathogen genome data. In preparation.

{{< bi file-earmark-pdf-fill >}} Di Tommaso, Paolo, Maria Chatzou, Evan W Floden, Pablo Prieto Barja, Emilio Palumbo, and Cedric Notredame. 2017. Nextflow Enables Reproducible Computational Workflows. Nature Biotechnology 35 (4): 316--19. https://doi.org/10.1038/nbt.3820.

Icons for this report were sampled from [Bootstrap Icons](https://icons.getbootstrap.com), [Freepick](https://www.freepik.com), [Academicons](https://jpswalsh.github.io/academicons/), and [Font Awesome](https://fontawesome.com).
